---
title: "Untitled"
output: html_document
---
## R Markdown

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Abstract 
## 1.Importing Libraries
## 1.1 Dataset Summary (description of dataset)
## 1.2 Data Cleaning & Quality 
## 2. EDA (Explanatory data analysis) - graphical analysis 
## 2.1 PCA + visual analysis 
## 2.2 Cluster Analysis 
----- Machine learning 
## Logistic regression
## Neural Networks 
## * Others etcc
---- -HPCI 
## Use R packages such as caret & parallel (to implement)
## Conclusion

```{r}
#Required libraries
library(ggplot2)
library(tidyverse)
library(dplyr)
library(data.table)
```

*---Research Question: Predict the energy assumption/power generation based on weather conditions---*


```{r}
#Required libraries
library(ggplot2)
library(tidyverse)
library(dplyr)
library(data.table)
```

*---Research Question: Predict the energy assumption/power generation based on weather conditions---*

```{r}
#Weather csv files 
weather.daily <- read.csv("weather_daily.csv")
weather.hourly <- read.csv("weather_hourly.csv")
```
 

```{r}
# Combined weather data set
weather.df <- merge(weather.daily, weather.hourly, by=c("time","visibility","windBearing","icon","dewPoint","windSpeed",'pressure',"summary","precipType","humidity"), all.x=TRUE,all.y=FALSE)

#Removing empty columns
weather2 = select(weather.df, -c(temperature,apparentTemperature))
```

```{r}
#Energy/power data sets 
energy.df <- read.csv("combined.csv")
weather2$time <- as.Date(weather2$time)
names(energy.df)[2] ="time"
energy.df$time <- as.Date(energy.df$time)
```


```{r}
#weather&energy
energyvibes<- inner_join(weather2, energy.df, by="time")
```























































****************************************************
```{r}
#Weather csv files 
weather.daily <- read.csv("weather_daily.csv")
weather.hourly <- read.csv("weather_hourly.csv")

#Power data set  (sample of the first 1000 rows)
power.head<- read.csv("power.1000lines.csv")
power.tails <- read.csv("power.tail.csv")
```


```{r}
#Renaming DateTime column in prep for combining 
names(power.head)[3]= "time"
names(power.head)[4]= 'energy'
names(power.tails)[1]= "LCLid"
names(power.tails)[2]= "stdorToU"
names(power.tails)[3]= "time"
names(power.tails)[4]= "energy"
names(power.tails)[5]= "Acorn"
names(power.tails)[6]= "Acorn_grouped"
```

```{r}
# Combining power data sets 
power.df<- merge(power.head, power.tails, by=c("LCLid", "time","energy","stdorToU","Acorn","Acorn_grouped"), all.x=TRUE, all.y=TRUE)

more.df <- read.csv("combined.csv")

power.df$time <- as.Date(power.df$time)

names(more.df)[2] = "time"


#Combined with more energy 
ultimate.power<- merge(power.df, more.df, by=c("LCLid", "time"),all.x=TRUE, all.y=TRUE)



#Combining weather data sets
weather.df <- merge(weather.daily, weather.hourly, by=c("time","visibility","windBearing","icon","dewPoint","windSpeed",'pressure',"summary","precipType","humidity"), all.x=TRUE,all.y=FALSE)

#Removing empty columns
weather2 = select(weather.df, -c(temperature,apparentTemperature))
```

-- I have noticed a difference in the amount of obs when you write (all.x=TRUE, all.y=False) vs (all.x=TRUE ,all.y=TRUE)
The code I ended up settling for has the same number of obs of weather.daily data set - quite strange but I am not sure where I went wrong there
Also it does come with less missing values'
```{r}
# Combining both weather and power data sets together
all.together <- merge(weather2, power.df, by="time", all.x=TRUE,all.y=TRUE)

colSums(is.na(all.together))
```


SO MANYY MISSING VALUES maybe i am merging it all wrong... 


/Users/munamohamed/Downloads/daily_dataset

```{r}
more.df <- read.csv("combined.csv")
```










