---
title: "CS5811_New"
output: html_document
---
## Abstract 
## 1.Importing Libraries
## 1.1 Dataset Summary (description of dataset)
## 1.2 Data Cleaning & Quality 
## 2. EDA (Explanatory data analysis) - graphical analysis 
## 2.1 PCA + visual analysis 
## 2.2 Cluster Analysis 
----- Machine learning 
## Logistic regression
## Neural Networks 
## * Others etcc
---- -HPCI 
## Use R packages such as caret & parallel (to implement)
## Conclusion


```{r}
#Required libraries
library(ggplot2)
library(tidyverse)
library(dplyr)
library(data.table)
```

*---Research Question: Predict the energy assumption/power generation based on weather conditions---*

```{r}
#Weather csv files 
weather.daily <- read.csv("weather_daily.csv")
weather.hourly <- read.csv("weather_hourly.csv")
```


```{r}
# Combined weather data set
weather.df <- merge(weather.daily, weather.hourly, by=c("time","visibility","windBearing","icon","dewPoint","windSpeed",'pressure',"summary","precipType","humidity"), all.x=TRUE,all.y=FALSE)

#Removing empty columns
weather2 = select(weather.df, -c(temperature,apparentTemperature, cloudCover, uvIndex, icon))
```

```{r}
#Energy/power data sets 
energy.df <- read.csv("combined.csv")
weather2$time <- as.Date(weather2$time)
names(energy.df)[2] ="time"
energy.df$time <- as.Date(energy.df$time)
```


```{r}
#weather&energy
energyvibes<- inner_join(weather2, energy.df, by="time")
```


```{r}
# Feature reduction

energy <- subset(energyvibes, select = -c(apparentTemperatureHighTime,
                                          apparentTemperatureLowTime, sunsetTime, sunriseTime,
                                          apparentTemperatureMinTime, uvIndexTime,apparentTemperatureHigh,
                                          apparentTemperatureLow, apparentTemperatureMaxTime, temperatureMaxTime,
                                          temperatureMinTime, apparentTemperatureMax, apparentTemperatureMin))

```

```{r}
summary(energy)
```

```{r}
energy$energy_median <- as.numeric(energy$energy_median)
energy$energy_mean <- as.numeric(energy$energy_mean)
energy$energy_max <- as.numeric(energy$energy_max)
energy$energy_count <- as.numeric(energy$energy_count)
energy$energy_std <- as.numeric(energy$energy_std)
energy$energy_sum <- as.numeric(energy$energy_sum)
energy$energy_min <- as.numeric(energy$energy_min)
```


```{r}
# Boxplots of all variables 
# This gives an insight into outliers compared to other variables
num_df<-energy%>%
  select(where(is.numeric))
boxplot(num_df)
# energy std and sum (max) and enerfy count (min) 
```
- In comparison to each other some have a lot of outliers whereas others don't, so the 3 variables that will have outliers removed are as follows: energy_count, _std and _sum 

```{r}
# removing energy_count outlier values (Min) 
boxplot(energy$energy_count)
# outlier detection from the boxplot 
#   note: the boxplot command returns a summary statistics object
#     this object can be assigned to a variable and inspected
energy_count_boxplot <- boxplot(energy$energy_count)
# inspect the boxplot statistics
#   note: the returened value is a list object
energy_count_boxplot
# outlier values can be extracted from this object
#   they are stored in the element 'out' of the list
energy_count_boxplot$out
# get the minimum value that is an outlier
min(energy_count_boxplot$out)
# outliers rows can be extracted by conditional selection
energy[energy$energy_count >= min(energy_count_boxplot$out), ]
```

-- Repeat the same code for (energy std & energy sum)

```{r}
# removing energy_sum outlier values (max) 
boxplot(energy$energy_sum)
#  this object can be assigned to a variable and inspected
energy_sum_boxplot <- boxplot(energy$energy_sum)
#   note: the returened value is a list object
energy_sum_boxplot
#   they are stored in the element 'out' of the list
energy_sum_boxplot$out
# get the max value that is an outlier
max(energy_sum_boxplot$out)
# outliers rows can be extracted by conditional selection
energy[energy$energy_sum >= max(energy_sum_boxplot$out), ]
```


```{r}
# removing energy_std outlier values (max) 
boxplot(energy$energy_std)
# outlier detection from the boxplot 
energy_std_boxplot <- boxplot(energy$energy_std)
# inspect the boxplot statistics
energy_std_boxplot
#   they are stored in the element 'out' of the list
energy_std_boxplot$out
# get the max value that is an outlier
max(energy_std_boxplot$out)
# outliers rows can be extracted by conditional selection
energy[energy$energy_std >= max(energy_std_boxplot$out), ]
```














