---
title: "Untitled"
output: html_document
---
## R Markdown

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
## Abstract 
## 1.Importing Libraries
## 1.1 Dataset Summary (description of dataset)
## 1.2 Data Cleaning & Quality 
## 2. EDA (Explanatory data analysis) - graphical analysis 
## 2.1 PCA + visual analysis 
## 2.2 Cluster Analysis 
----- Machine learning 
## Logistic regression
## Neural Networks 
## * Others etcc
---- -HPCI 
## Use R packages such as caret & parallel (to implement)
## Conclusion


```{r}
#Required libraries
library(ggplot2)
library(tidyverse)
library(dplyr)
library(data.table)
```

*---Research Question: Predict the energy assumption/power generation based on weather conditions---*

```{r}
#Weather csv files 
weather.daily <- read.csv("weather_daily.csv")
weather.hourly <- read.csv("weather_hourly.csv")
```
 

```{r}
# Combined weather data set
weather.df <- merge(weather.daily, weather.hourly, by=c("time","visibility","windBearing","icon","dewPoint","windSpeed",'pressure',"summary","precipType","humidity"), all.x=TRUE,all.y=FALSE)

#Removing empty columns
weather2 = select(weather.df, -c(temperature,apparentTemperature, cloudCover, uvIndex, icon))
```

```{r}
#Energy/power data sets 
energy.df <- read.csv("combined.csv")
weather2$time <- as.Date(weather2$time)
names(energy.df)[2] ="time"
energy.df$time <- as.Date(energy.df$time)
```


```{r}
#weather&energy
energyvibes<- inner_join(weather2, energy.df, by="time")
```


```{r}
# Feature reduction

energy <- subset(energyvibes, select = -c(apparentTemperatureHighTime,
                                          apparentTemperatureLowTime, sunsetTime, sunriseTime,
                                          apparentTemperatureMinTime, uvIndexTime,apparentTemperatureHigh,
                                          apparentTemperatureLow, apparentTemperatureMaxTime, temperatureMaxTime,
                                          temperatureMinTime, apparentTemperatureMax, apparentTemperatureMin))

```

```{r}
summary(energy)
```

```{r}
energy$energy_median <- as.numeric(energy$energy_median)
energy$energy_mean <- as.numeric(energy$energy_mean)
energy$energy_max <- as.numeric(energy$energy_max)
energy$energy_count <- as.numeric(energy$energy_count)
energy$energy_std <- as.numeric(energy$energy_std)
energy$energy_sum <- as.numeric(energy$energy_sum)
energy$energy_min <- as.numeric(energy$energy_min)
```

```{r}
# Checking outliers
boxplot(energy$energy_median)
boxplot(energy$energy_mean)
boxplot(energy$energy_max)
boxplot(energy$energy_count)
boxplot(energy$energy_std)
boxplot(energy$energy_sum)
boxplot(energy$energy_min)
```

```{r}
boxplot(energy$visibility)
boxplot(energy$windBearing)
boxplot(energy$dewPoint)
boxplot(energy$windSpeed)
boxplot(energy$pressure)
boxplot(energy$humidity)
boxplot(energy$temperatureMax)
boxplot(energy$temperatureLow)
boxplot(energy$temperatureMin)
boxplot(energy$temperatureHigh)
boxplot(energy$moonPhase)
```

```{r}
box_median <- boxplot(energy$energy_median, plot = F)$out
d1 <- energy[!(energy$energy_median %in% box_median), ]
boxplot(d1$energy_median)
```

```{r}
box_mean <- boxplot(d1$energy_mean, plot = F)$out
d2 <- d1[!(d1$energy_mean %in% box_mean), ]
boxplot(d2$energy_mean)
```

```{r}
box_max <- boxplot(d2$energy_max, plot = F)$out
d3 <- d2[!(d2$energy_max %in% box_max), ]
boxplot(d3$energy_max)
```

# energy_count has a weird boxplot


```{r}
box_std <- boxplot(d3$energy_std, plot = F)$out
d5 <- d3[!(d3$energy_std %in% box_std), ]
boxplot(d5$energy_std)
```

```{r}
box_sum <- boxplot(d5$energy_sum, plot = F)$out
d6 <- d5[!(d5$energy_sum %in% box_sum), ]
boxplot(d6$energy_sum)
```

```{r}
box_min <- boxplot(d6$energy_min, plot = F)$out
d7 <- d6[!(d6$energy_min %in% box_min), ]
boxplot(d7$energy_min)
```

```{r}
box_visibility <- boxplot(d7$visibility, plot = F)$out
d8 <- d7[!(d7$visibility %in% box_visibility), ]
boxplot(d8$visibility)
```

```{r}
box_windspeed <- boxplot(d8$windSpeed, plot = F)$out
d9 <- d8[!(d8$windSpeed %in% box_windspeed), ]
boxplot(d8$windSpeed)
```

```{r}
box_pressure <- boxplot(d9$pressure, plot = F)$out
d10 <- d9[!(d8$pressure %in% box_pressure), ]
boxplot(d10$pressure)
```

```{r}
box_humidity <- boxplot(d10$humidity, plot = F)$out
d11 <- d10[!(d10$humidity %in% box_humidity), ]
boxplot(d10$humidity)
```

```{r}
sum(is.na(d11))
```

```{r}
colSums(is.na(d11))
```

```{r}
# Removing observations with missing values
data <- na.omit(d11)
```

```{r}
set.seed(100)
# Subsampling the data
data_clean <- data[sample(nrow(data), 100000, replace = FALSE),]
```

```{r}
summary(data_clean)
```

```{r}
data_clean$precipType <- as.factor(data_clean$precipType)
```

```{r}
sum(duplicated(data_clean))
```

```{r}
write.csv(data_clean, "/Users/RoshniPatel/Documents/data.csv", row.names = FALSE)
```


****************************************************








