---
title: "Untitled"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## R Markdown

## Abstract 
## 1.Importing Libraries
## 1.1 Dataset Summary (description of dataset)
## 1.2 Data Cleaning & Quality 
## 2. EDA (Explanatory data analysis) - graphical analysis 
## 2.1 PCA + visual analysis 
## 2.2 Cluster Analysis 

----- Machine learning 
## Logistic regression
## Neural Networks 
## * Others etcc

---- -HPCI 
## Use R packages such as caret & parallel (to implement)
## Conclusion

```{r}
library(ggplot2)
library(tidyverse)
library(dplyr)
```

***CLEAN WEATHER FIRST THEN COMBINEEEE!!!!!!!!!!!!!!***

---Research Question: Predict the energy assumption/power generation based on weather conditions---

*Loading required packages*
```{r}
#Weather csv files 
weather.hour <- read.csv("weather_hourly.csv")
weather.daily <- read.csv("weather_daily.csv")

#Combined weather datasets by time 
weather <-merge(weather.hour, weather.daily, by.x ="time", by.y="time", all.x= TRUE, all.y= FALSE)
```


```{r}
str(weather)
```

```{r}
colSums(is.na(weather))
```

```{r}
#Power data set  (sample of the first 1000 rows)
power <- read.csv("power.1000lines.csv")

#Renaming DateTime column in prep for combining 
names(power)[3]= "time"
```


```{r}
# Combined data sets 
combined = full_join(weather,power, by ="time")
# Changing the time column to just time and not hours too
combined$time =  as.Date(combined$time)

colSums(is.na(combined))

```






```{r}
#Merging the power files together
combined <-merge(power.head.sample, weather, by.x ="DateTime", by.y="time", all.x= TRUE, all.y= FALSE)

combined$DateTime <- as.Date(combined$DateTime)
```















