---
  title: "CS5811_New"
output: html_document
---
  ## Abstract 
  ## 1.Importing Libraries
  ## 1.1 Dataset Summary (description of dataset)
  ## 1.2 Data Cleaning & Quality 
  ## 2. EDA (Explanatory data analysis) - graphical analysis 
  ## 2.1 PCA + visual analysis 
  ## 2.2 Cluster Analysis 
  ----- Machine learning 
## Logistic regression
## Neural Networks 
## * Others etcc
---- -HPCI 
## Use R packages such as caret & parallel (to implement)
## Conclusion


```{r}
#Required libraries
library(ggplot2)
library(tidyverse)
library(dplyr)
library(data.table)
library(lubridate)
library(psych)
library(scales)
```

##---Research Question: Predict the energy consumption/power generation based on weather conditions

```{r}
#Weather csv files 
weather.daily <- read.csv("weather_daily.csv")
weather.hourly <- read.csv("weather_hourly.csv")
```

```{r}
head(weather.daily)
head(weather.hourly)
```

```{r}
# Combining the two data sets weather_daily and weather_hourly
weather.df <- merge(weather.daily, weather.hourly, by=c("time","visibility","windBearing","icon","dewPoint","windSpeed",'pressure',"summary","precipType","humidity"), all.x=TRUE,all.y=FALSE)
```

```{r}
#Checking and removing empty columns and duplicating time column as time and date
weather2 = select(weather.df, -c(temperature,apparentTemperature, cloudCover, uvIndex, icon))
weather2 <- weather2 %>% mutate(timeanddate = time)
```

```{r}
#cleaning empty spaces between and after the data
trimws(weather2$time)
trimws(weather2$icon)
trimws(weather2$summary)
trimws(weather2$precipType)
trimws(weather2$temperatureMaxTime)
trimws(weather2$temperatureMinTime)
trimws(weather2$apparentTemperatureMinTime)
trimws(weather2$apparentTemperatureHighTime)
trimws(weather2$sunsetTime)
trimws(weather2$sunriseTime)
trimws(weather2$temperatureHighTime)
trimws(weather2$uvIndexTime)
trimws(weather2$temperatureLowTime)
trimws(weather2$apparentTemperatureMaxTime)
trimws(weather2$apparentTemperatureLowTime)
```


```{r}
#changing date to POSIXct
#showing 00:00:00 for midnight 
print.POSIXct <- function(x,...)print(format(x,"%Y-%m-%d %H:%M:%S"))

weather2$time               <- as.POSIXct(strptime(weather2$time, format = "%d/%m/%Y %H:%M",tz="GMT"))

weather2$weatherDay         <- format(as.POSIXct(weather2$time,format="%Y/%m/%d %H:%M:%S"),"%Y/%m/%d")

weather2$timeanddate        <- as.POSIXct(strptime(weather2$timeanddate,
                                                   format = "%d/%m/%Y %H:%M",tz="GMT"))

weather2$temperatureMaxTime <- as.POSIXct(strptime (weather2$temperatureMaxTime,
                                                    format = "%d/%m/%Y %H:%M",tz="GMT"))

weather2$temperatureMinTime <- as.POSIXct(strptime (weather2$temperatureMinTime,
                                                    format = "%d/%m/%Y %H:%M",tz="GMT"))

weather2$apparentTemperatureMinTime   <- as.POSIXct(strptime (weather2$apparentTemperatureMinTime,
                                                              format = "%d/%m/%Y %H:%M",tz="GMT"))

weather2$apparentTemperatureHighTime  <- as.POSIXct(strptime (weather2$apparentTemperatureHighTime,
                                                              format = "%d/%m/%Y %H:%M",tz="GMT"))

weather2$sunsetTime         <- as.POSIXct(strptime(weather2$sunsetTime,
                                                   format = "%d/%m/%Y %H:%M",tz="GMT"))

weather2$sunriseTime        <- as.POSIXct(strptime(weather2$sunriseTime,
                                                   format = "%d/%m/%Y %H:%M",tz="GMT"))

weather2$temperatureHighTime<- as.POSIXct(strptime(weather2$temperatureHighTime,
                                                   format = "%d/%m/%Y %H:%M",tz="GMT"))

weather2$uvIndexTime        <- as.POSIXct(strptime(weather2$uvIndexTime,
                                                   format = "%d/%m/%Y %H:%M",tz="GMT"))

weather2$temperatureLowTime <- as.POSIXct(strptime(weather2$temperatureLowTime,
                                                   format = "%d/%m/%Y %H:%M",tz="GMT"))
 
weather2$apparentTemperatureMaxTime   <- as.POSIXct(strptime
                                                    (weather2$apparentTemperatureMaxTime,
                                                      format = "%d/%m/%Y %H:%M",tz="GMT"))

weather2$apparentTemperatureLowTime   <- as.POSIXct(strptime
                                                    (weather2$apparentTemperatureLowTime,
                                                      format = "%d/%m/%Y %H:%M",tz="GMT"))

weather2$windBearing    <- as.numeric(weather2$windBearing)
weather2$precipType     <- as.factor(weather2$precipType)
weather2$summary        <- as.factor(weather2$summary)

```

```{r}
str(weather2)
head(weather2)
tail(weather2)
```
```{r}
weather3 <- weather2
```

```{r}
#Energy/power data sets 
energy.df <- read.csv("combined.csv")
#names(energy.df)[2] ="time"
```

```{r}
str(energy.df)
```

```{r}
#converting data to numerical and char to date

energy.df$day           <- as.Date(energy.df$day, format = "%d/%m/%Y")
energy.df$LCLid         <- as.factor(energy.df$LCLid)
energy.df$energy_median <- as.numeric(energy.df$energy_median)
energy.df$energy_mean   <- as.numeric(energy.df$energy_mean)
energy.df$energy_max    <- as.numeric(energy.df$energy_max)
energy.df$energy_count  <- as.numeric(energy.df$energy_count)
energy.df$energy_std    <- as.numeric(energy.df$energy_std)
energy.df$energy_sum    <- as.numeric(energy.df$energy_sum)
energy.df$energy_min    <- as.numeric(energy.df$energy_min)
```

```{r}
# Checking converted data
str(energy.df)
head(energy.df)
head(energy.df)
```

```{r}
#creating column to join Weather2 and energy.df with weather2 column  weatherDay and day from energy.df

weather2$weatherDay <- as.Date(weather2$weatherDay)
head(weather2$weatherDay)
tail(weather2$weatherDay)
```
```{r}
#weather&energy

energyvibes<- inner_join(energy.df, weather2, by = c("day" = "weatherDay"))
```

```{r}
#checking joined data
head(energyvibes)
tail(energyvibes)
str(energyvibes)
```

```{r}
#backup of All combined data
#write.csv(energyvibes, "energyvibesDatamerged.csv")
```

```{r}
# Feature reduction

#energy <- subset(energyvibes, select = -c(apparentTemperatureHighTime,
#                                          apparentTemperatureLowTime, sunsetTime, sunriseTime,
#                                          apparentTemperatureMinTime,
#                                          uvIndexTime,apparentTemperatureHigh,
#                                          apparentTemperatureLow, apparentTemperatureMaxTime,
#                                          temperatureMaxTime,
#                                          temperatureMinTime, apparentTemperatureMax,
#                                          apparentTemperatureMin))


energy <- subset(energyvibes, select = -c(apparentTemperatureHighTime,
                                          apparentTemperatureLowTime,
                                          apparentTemperatureMinTime,
                                          uvIndexTime,
                                          apparentTemperatureHigh,
                                          apparentTemperatureLow,
                                          apparentTemperatureMaxTime,
                                          apparentTemperatureMax,
                                          apparentTemperatureMin,
                                          temperatureMinTime,
                                          temperatureMaxTime))
```


```{r}
summary(energy)
str(energy)
head(energy)
```

```{r}
#checking uniqueness of date w.r.t time and household id for uniqueness
energyHouseIdDuplicate      <- energy[,c('LCLid','time')]
energyHouseIdDuplicateCheck <- energy[duplicated(energyHouseIdDuplicate) |
                                        duplicated(energyHouseIdDuplicate, fromLast=TRUE),]

energyHouseIdDuplicateCheck
#no duplicates found

#checking duplicates w.r.t household
energyHouseIdDuplicate      <- energy[,c('LCLid','day')]
energyHouseIdDuplicateCheck <- energy[duplicated(energyHouseIdDuplicate) |
                                        duplicated(energyHouseIdDuplicate, fromLast=TRUE),]
energyHouseIdDuplicateCheck
#4072 rows resulted in duplicates

#removing duplicates
energyUnique   <- energy %>% distinct(energy$LCLid, energy$day, .keep_all = TRUE)
energyUnique

#rechecking duplicate
#checking duplicates w.r.t household

energyHouseIdDuplicateRecheck <- energyUnique[,c('LCLid','day')]
energyHouseIdDuplicateRecheck <- energyUnique[duplicated(energyHouseIdDuplicateRecheck) |
                                        duplicated(energyHouseIdDuplicateRecheck, fromLast=TRUE),]

energyHouseIdDuplicateRecheck
# 0 results for duplicates
```
```{r}
summary(energyUnique)
str(energyUnique)
```
```{r}
#data for work
#write.csv(energyUnique, "energyData.csv")
```
```{r}
#creating sub set of data
energy_weather_data <- sample_n(energyUnique, 1000)
head(energy_weather_data)
tail(energy_weather_data)
nrow(energy_weather_data)
```

#breaking down time column in
```{r}
#breaking down of time column
energy_weather_data$energy_time_day   <- format(energy_weather_data$time, format = "%d")
energy_weather_data$energy_time_month <- format(energy_weather_data$time, format = "%m")
energy_weather_data$energy_time_year  <- format(energy_weather_data$time, format = "%Y")
energy_weather_data$energy_time_hour  <- format(energy_weather_data$time, format = "%H")
energy_weather_data$energy_time_min   <- format(energy_weather_data$time, format = "%M")

#breaking down of temperature temperatureHighTime column
energy_weather_data$temperature_Hightime_day   <- format(energy_weather_data$temperatureHighTime,
                                                         format = "%d") 
energy_weather_data$temperature_Hightime_month <- format(energy_weather_data$temperatureHighTime,
                                                         format = "%m")
energy_weather_data$temperature_Hightime_year  <- format(energy_weather_data$temperatureHighTime,
                                                         format = "%Y")
energy_weather_data$temperature_Hightime_hour  <- format(energy_weather_data$temperatureHighTime,
                                                         format = "%H")
energy_weather_data$temperature_Hightime_min   <- format(energy_weather_data$temperatureHighTime,
                                                         format = "%M")

#breaking down of temperature temperatureLowTime column
energy_weather_data$temperature_Lowtime_day   <- format(energy_weather_data$temperatureLowTime,
                                                         format = "%d") 
energy_weather_data$temperature_Lowtime_month <- format(energy_weather_data$temperatureLowTime,
                                                         format = "%m")
energy_weather_data$temperature_Lowtime_year  <- format(energy_weather_data$temperatureLowTime,
                                                         format = "%Y")
energy_weather_data$temperature_Lowtime_hour  <- format(energy_weather_data$temperatureLowTime,
                                                         format = "%H")
energy_weather_data$temperature_Lowtime_min   <- format(energy_weather_data$temperatureLowTime,
                                                         format = "%M")

#breaking down of temperature sunriseTime column
energy_weather_data$sunriseTime_day   <- format(energy_weather_data$sunriseTime,
                                                         format = "%d") 
energy_weather_data$sunriseTime_month <- format(energy_weather_data$sunriseTime,
                                                         format = "%m")
energy_weather_data$sunriseTime_year  <- format(energy_weather_data$sunriseTime,
                                                         format = "%Y")
energy_weather_data$sunriseTime_hour  <- format(energy_weather_data$sunriseTime,
                                                         format = "%H")
energy_weather_data$sunriseTime_min   <- format(energy_weather_data$sunriseTime,
                                                         format = "%M")

#breaking down of temperature sunsetTime column
energy_weather_data$sunsetTime_day   <- format(energy_weather_data$sunsetTime,
                                                         format = "%d") 
energy_weather_data$sunsetTime_month <- format(energy_weather_data$sunsetTime,
                                                         format = "%m")
energy_weather_data$sunsetTime_year  <- format(energy_weather_data$sunsetTime,
                                                         format = "%Y")
energy_weather_data$sunsetTime_hour  <- format(energy_weather_data$sunsetTime,
                                                         format = "%H")
energy_weather_data$sunsetTime_min   <- format(energy_weather_data$sunsetTime,
                                                         format = "%M")
head(energy_weather_data)
```


```{r}
#converting newly formed deisintegrated time and date char to num
energy_weather_data$energy_time_day   <- as.numeric(energy_weather_data$energy_time_day)
energy_weather_data$energy_time_month <- as.numeric(energy_weather_data$energy_time_month)
energy_weather_data$energy_time_year  <- as.numeric(energy_weather_data$energy_time_year)
energy_weather_data$energy_time_hour  <- as.numeric(energy_weather_data$energy_time_hour)
energy_weather_data$energy_time_min   <- as.numeric(energy_weather_data$energy_time_min)

energy_weather_data$temperature_Hightime_day   <- as.numeric(
                                                  energy_weather_data$temperature_Hightime_day)
energy_weather_data$temperature_Hightime_month <- as.numeric(
                                                  energy_weather_data$temperature_Hightime_month)
energy_weather_data$temperature_Hightime_year  <- as.numeric(
                                                  energy_weather_data$temperature_Hightime_year)
energy_weather_data$temperature_Hightime_hour  <- as.numeric(
                                                  energy_weather_data$temperature_Hightime_hour)
energy_weather_data$temperature_Hightime_min   <- as.numeric(
                                                  energy_weather_data$temperature_Hightime_min)

energy_weather_data$temperature_Lowtime_day    <- as.numeric(
                                                  energy_weather_data$temperature_Lowtime_day)
energy_weather_data$temperature_Lowtime_month  <- as.numeric(
                                                  energy_weather_data$temperature_Lowtime_month)
energy_weather_data$temperature_Lowtime_year   <- as.numeric(
                                                  energy_weather_data$temperature_Lowtime_year)
energy_weather_data$temperature_Lowtime_hour   <- as.numeric(
                                                  energy_weather_data$temperature_Lowtime_hour)
energy_weather_data$temperature_Lowtime_min    <- as.numeric(
                                                  energy_weather_data$temperature_Lowtime_min)

energy_weather_data$sunriseTime_day   <- as.numeric(energy_weather_data$sunriseTime_day)
energy_weather_data$sunriseTime_month <- as.numeric(energy_weather_data$sunriseTime_month)
energy_weather_data$sunriseTime_year  <- as.numeric(energy_weather_data$sunriseTime_year)
energy_weather_data$sunriseTime_hour  <- as.numeric(energy_weather_data$sunriseTime_hour)
energy_weather_data$sunriseTime_min   <- as.numeric(energy_weather_data$sunriseTime_min)

energy_weather_data$sunsetTime_day    <- as.numeric(energy_weather_data$sunsetTime_day)
energy_weather_data$sunsetTime_month  <- as.numeric(energy_weather_data$sunsetTime_month)
energy_weather_data$sunsetTime_year   <- as.numeric(energy_weather_data$sunsetTime_year)
energy_weather_data$sunsetTime_hour   <- as.numeric(energy_weather_data$sunsetTime_hour)
energy_weather_data$sunsetTime_min    <- as.numeric(energy_weather_data$sunsetTime_min)
```

```{r}
str(energy_weather_data)
```
```{r}
#checking Na's occurrence in the data set energy_weather_data 
weather_dataset_NA_count <- apply(is.na(energy_weather_data),2,sum)
weather_dataset_NA_count

# one occurrence found in energy_std imputing
median_energy_std <- median(energy_weather_data$energy_std, na.rm = T)
energy_weather_data[is.na(energy_weather_data$energy_std), 'energy_std'] = median_energy_std
```


```{r}
#boxplots to see the outliers of columns
boxplot(energy_weather_data$time, main = "Energy Time", xlab = "time and date", ylab = "time and date",
        col = "green", border = "yellow", horizontal = TRUE, notch = TRUE,origin)

boxplot(energy_weather_data$sunsetTime, main = "Sun Set Time",
        xlab = "Sun Set Time", ylab = "Sun Set Time", 
        col = "green", border = "yellow", horizontal = TRUE, notch = TRUE)

boxplot(energy_weather_data$sunriseTime, main = "Sun Rise Time",
        xlab = "Sun Rise Time", ylab = "Sun Rise Time", 
        col = "green", border = "yellow", horizontal = TRUE, notch = TRUE)

boxplot(energy_weather_data$temperatureHighTime, main = "High Temperatre Time",
        xlab = "High Temperatre Time", ylab = "High Temperatre Time", 
        col = "green", border = "yellow", horizontal = TRUE, notch = TRUE)

boxplot(energy_weather_data$temperatureLowTime, main = "Low Temperatre Time",
        xlab = "Low Temperatre Time", ylab = "Low Temperatre Time", 
        col = "green", border = "yellow", horizontal = TRUE, notch = TRUE)

boxplot(energy_weather_data$timeanddate, main = "Time and Date",
        xlab = "Time and Date", ylab = "Time and Date", 
        col = "green", border = "yellow", horizontal = TRUE, notch = TRUE)

```

```{r}
#boxplots to see the outliers of columns
boxplot(energy_weather_data$visibility, main = "Visibility", xlab = "Visibility", ylab = "Visibility",
        col = "green", border = "yellow", horizontal = TRUE, notch = TRUE)

boxplot(energy_weather_data$windBearing, main = "Wind Bearing", xlab = "Wind Bearing", ylab = "Wind Bearing",
        col = "green", border = "yellow", horizontal = TRUE, notch = TRUE)

boxplot(energy_weather_data$dewPoint, main = "Dew Point", xlab = "Dew Point", ylab = "Dew Point",
        col = "green", border = "yellow", horizontal = TRUE, notch = TRUE)


boxplot(energy_weather_data$windSpeed, main = "Wind Speed", xlab = "Wind Speed", ylab = "Wind Speed",
        col = "green", border = "yellow", horizontal = TRUE, notch = TRUE)

boxplot(energy_weather_data$pressure, main = "Pressure", xlab = "Pressure", ylab = "Pressure",
        col = "green", border = "yellow", horizontal = TRUE, notch = TRUE)

boxplot(energy_weather_data$humidity, main = "Humidity", xlab = "Humidity", ylab = "Humidity",
        col = "green", border = "yellow", horizontal = TRUE, notch = TRUE)

boxplot(energy_weather_data$temperatureMax, main = "Maximum Temperature", xlab = "Maximum Temperature",
        ylab = "Maximum Temperature", col = "green", border = "yellow", horizontal = TRUE, notch = TRUE)

boxplot(energy_weather_data$temperatureLow, main = "Low Temperature Time",
        xlab = "Low Temperature Time", ylab = "Low Temperature Time", 
        col = "green", border = "yellow", horizontal = TRUE, notch = TRUE)

boxplot(energy_weather_data$temperatureMin, main = "Minimum Temperature",
        xlab = "Minimum Temperature", ylab = "Minimum Temperature", 
        col = "green", border = "yellow", horizontal = TRUE, notch = TRUE)

boxplot(energy_weather_data$temperatureHigh, main = "Maximum Temperature",
        xlab = "Maximum Temperature", ylab = "Maximum Temperature", 
        col = "green", border = "yellow", horizontal = TRUE, notch = TRUE)

boxplot(energy_weather_data$moonPhase, main = "Moon Phase",
        xlab = "Moon Phase", ylab = "Moon Phase", 
        col = "green", border = "yellow", horizontal = TRUE, notch = TRUE)

boxplot(energy_weather_data$energy_median, main = "Energy Median",
        xlab = "Energy Median", ylab = "Energy Median", 
        col = "green", border = "yellow", horizontal = TRUE, notch = TRUE)

boxplot(energy_weather_data$energy_mean, main = "Energy Mean",
        xlab = "Energy Mean", ylab = "Energy Mean", 
        col = "green", border = "yellow", horizontal = TRUE, notch = TRUE)

boxplot(energy_weather_data$energy_max, main = "Energy Maximum",
        xlab = "Energy Maximum", ylab = "Energy Maximum", 
        col = "green", border = "yellow", horizontal = TRUE, notch = TRUE)

boxplot(energy_weather_data$energy_count, main = "Energy Count",
        xlab = "Energy Count", ylab = "Energy Count", 
        col = "green", border = "yellow", horizontal = TRUE, notch = TRUE)

boxplot(energy_weather_data$energy_std, main = "Energy STD",
        xlab = "Energy STD", ylab = "Energy STD", 
        col = "green", border = "yellow", horizontal = TRUE, notch = TRUE)

boxplot(energy_weather_data$energy_sum, main = "Energy Sum",
        xlab = "Energy Sum", ylab = "Energy Sum", 
        col = "green", border = "yellow", horizontal = TRUE, notch = TRUE)

boxplot(energy_weather_data$energy_min, main = "Energy Minimum",
        xlab = "Energy Minimum", ylab = "Energy Minimum", 
        col = "green", border = "yellow", horizontal = TRUE, notch = TRUE)

#columns with outliers
#visibility-, windSpeed+,  pressure-, humidity-, energy_median+, energy_mean+, energy_max+, energy_count-, energy_std+, energy_sum+, energy_min+
```

```{r}
#removing outliers from the data set energy_weather_data **
weather_dataset_boxplot <- boxplot(energy_weather_data$visibility)
max_visibility_outlier  <- max(weather_dataset_boxplot$out)
test1 <- energy_weather_data[energy_weather_data$visibility < max_visibility_outlier, ]
```
```{r}
#relationship between time and max energy consumption
ggplot(energy_weather_data, aes(time, energy_max)) +
       geom_point() +
       theme(axis.text.x = element_text(angle = 90, hjust = 1)) +
       scale_x_datetime(labels = date_format("%d:%m:%Y"))
```

## Removing outliers
```{r}
# Boxplots of all variables 
# This gives an insight into outliers compared to numerical variables
num_df <- energyUnique%>% select(where(is.numeric))
boxplot(num_df)
```




```{r}

#Outliers energy_median
```
# In comparison to each other some have a lot of outliers whereas others don't, so the 3 variables that will have outliers removed are as follows: energy_count, _std and _sum 

```{r}
# removing energy_count outlier values (Min) 
boxplot(energy$energy_count)
# outlier detection from the boxplot 
#   note: the boxplot command returns a summary statistics object
#     this object can be assigned to a variable and inspected
energy_count_boxplot <- boxplot(energy$energy_count)
# inspect the boxplot statistics
#   note: the returened value is a list object
energy_count_boxplot
# outlier values can be extracted from this object
#   they are stored in the element 'out' of the list
energy_count_boxplot$out
# get the minimum value that is an outlier
min(energy_count_boxplot$out)
# outliers rows can be extracted by conditional selection
energy[energy$energy_count >= min(energy_count_boxplot$out), ]
```

-- Repeat the same code for (energy std & energy sum)

```{r}
# removing energy_sum outlier values (max) 
boxplot(energy$energy_sum)
#  this object can be assigned to a variable and inspected
energy_sum_boxplot <- boxplot(energy$energy_sum)
#   note: the returened value is a list object
energy_sum_boxplot
#   they are stored in the element 'out' of the list
energy_sum_boxplot$out
# get the max value that is an outlier
max(energy_sum_boxplot$out)
# outliers rows can be extracted by conditional selection
energy[energy$energy_sum >= max(energy_sum_boxplot$out), ]
```


```{r}
# removing energy_std outlier values (max) 
boxplot(energy$energy_std)
# outlier detection from the boxplot 
energy_std_boxplot <- boxplot(energy$energy_std)
# inspect the boxplot statistics
energy_std_boxplot
#   they are stored in the element 'out' of the list
energy_std_boxplot$out
# get the max value that is an outlier
max(energy_std_boxplot$out)
# outliers rows can be extracted by conditional selection
energy[energy$energy_std >= max(energy_std_boxplot$out), ]
```


# Removing duplicate 
```{r}
# remove duplicate rows
energy_nodupes <- unique(energy)

# check the size of the data frames before and after removal
dim(energy)
dim(energy_nodupes)

# find the duplicate rows
#   duplicated returns a boolean vector
energy_duplicated <- duplicated(energy_nodupes)
energy_duplicated

# duplicated rows can then be extracted by conditional filtering
energy_nodupes[energy_duplicated, ]
```
# Removing missing values 


```{r}