---
title: "Untitled"
output: html_document
---
## Abstract 
## 1.Importing Libraries
## 1.1 Dataset Summary (description of dataset)
## 1.2 Data Cleaning & Quality 
## 2. EDA (Explanatory data analysis) - graphical analysis 
## 2.1 PCA + visual analysis 
## 2.2 Cluster Analysis 
----- Machine learning 
## Logistic regression
## Neural Networks 
## * Others etcc
---- -HPCI 
## Use R packages such as caret & parallel (to implement)
## Conclusion

------

## 1.Importing Libraries

```{r}
#Required libraries

library(ggplot2)
library(tidyverse)
library(dplyr)
library(data.table)
library(lubridate)
library(psych)
library(scales)
library("ggstatsplot")
library("rstantools")
library(wordcloud)  # word-cloud generator 
library(RColorBrewer) # color palettes3
library(tm)  # for text mining
```

                              ##---Research Question: Predict the energy consumption/power generation based on weather conditions


##  Importing Datasets 

```{r}
#Weather csv files 

weather.daily <- read.csv("weather_daily.csv")
weather.hourly <- read.csv("weather_hourly.csv")

#Energy/power data sets 
energy.df <- read.csv("combined.csv")
```



## Joining Weather datasets 

```{r}
# Combined weather data set
weather.df <- merge(weather.daily, weather.hourly, by=c("time","visibility","windBearing","icon","dewPoint","windSpeed",'pressure',"summary","precipType","humidity"), all.x=TRUE,all.y=FALSE)

#Removing empty columns
weather2 = select(weather.df, -c(temperature,apparentTemperature, cloudCover, uvIndex, icon))
```

```{r}
#Energy/power data sets 

weather2$time <- as.Date(weather2$time)
names(energy.df)[2] ="time"
energy.df$time <- as.Date(energy.df$time)
```


## Joining all datasets 

```{r}
#Combined weather&energy dataset 

energyvibes<- inner_join(weather2, energy.df, by="time")
```




## Data Cleaning 

```{r}
# Feature reduction

energy <- subset(energyvibes, select = -c(apparentTemperatureHighTime,
                                          apparentTemperatureLowTime, sunsetTime, sunriseTime,
                                         apparentTemperatureMinTime,
                                          uvIndexTime,apparentTemperatureHigh,
                                          apparentTemperatureLow, apparentTemperatureMaxTime,
                                          temperatureMaxTime, 
                                          temperatureMinTime, apparentTemperatureMax,
                                         apparentTemperatureMin))
```


```{r}
#removing variables from memory

rm(weather.daily)
rm(weather.hourly)
rm(weather.df)
rm(weather2)
rm(energy.df)
rm(energyvibes)
```

```{r}
# Changing data type

energy$windBearing <- as.numeric(energy$windBearing)
energy$LCLid         <- as.factor(energy$LCLid)
energy$energy_median <- as.numeric(energy$energy_median)
energy$energy_mean   <- as.numeric(energy$energy_mean)
energy$energy_max    <- as.numeric(energy$energy_max)
energy$energy_count  <- as.numeric(energy$energy_count)
energy$energy_std    <- as.numeric(energy$energy_std)
energy$energy_sum    <- as.numeric(energy$energy_sum)
energy$energy_min    <- as.numeric(energy$energy_min)
```


## Imputing missing values 

```{r}
#Checking the the no.of columns with missing values 

colSums(is.na(energy))

# The following variables have missing values:
# energy_median, mean,max,std,sum and min.
```

```{r}
# imputing missing values by replacing them with the median 
median_energy <- median(energy$energy_median, na.rm = T)
mean_energy <- median(energy$energy_mean, na.rm = T)
max_energy <- median(energy$energy_max, na.rm = T)
min_energy <- median(energy$energy_min, na.rm = T)
std_energy <- median(energy$energy_std, na.rm = T)
sum_energy <- median(energy$energy_sum, na.rm = T)

# once the median is discovered, remove the missing value by replacing it with the median 
energy[is.na(energy$energy_median), 'energy_median'] = median_energy
energy[is.na(energy$energy_mean), 'energy_mean'] = mean_energy
energy[is.na(energy$energy_max), 'energy_max'] = max_energy
energy[is.na(energy$energy_min), 'energy_min'] = min_energy
energy[is.na(energy$energy_std), 'energy_std'] = std_energy
energy[is.na(energy$energy_sum), 'energy_sum'] = sum_energy

# checking the dimension and if there are any missing values
dim(energy)
colSums(is.na(energy))
```

```{r}
energy$energy_median <- as.numeric(energy$energy_median)
```



## Removing outliers

```{r}
## Removing outliers
# Boxplots of all variables 
# This gives an insight into outliers compared to numerical variables

num_df <- energy %>% select(where(is.numeric)) %>% map(~ boxplot.stats(.x)$out)
par(mar=c(7,5,1,1))
boxplot(num_df, main = "Outliers", col = "green", border = "black", las = 2 )

# outliers: pressure, windSpeed, energy_sum,mean, std,max and min
```

```{r}
#imputing the outliers
#1 energy_median
median_weather_energy = boxplot(energy$energy_median)
min_median_energy     = min(median_weather_energy$out)
energy_weather_data_outliers = energy[energy$energy_median <
                                                      min_median_energy, ]

#2 energy_max
max_weather_energy = boxplot(energy_weather_data_outliers$energy_max)
min_max_energy     = min(max_weather_energy$out)
energy_weather_data_outliers = energy_weather_data_outliers[energy_weather_data_outliers$energy_max <
                                                        min_max_energy, ]

#3 energy_std
std_weather_energy = boxplot(energy_weather_data_outliers$energy_std)
min_std_energy     = min(std_weather_energy$out)
energy_weather_data_outliers = energy_weather_data_outliers[energy_weather_data_outliers$energy_std <
                                                        min_std_energy, ]
#4 energy_min
min_weather_energy  = boxplot(energy_weather_data_outliers$energy_min)
min_min_energy      = min(min_weather_energy$out)
energy_weather_data_outliers = energy_weather_data_outliers[energy_weather_data_outliers$energy_min <
                                                       min_min_energy, ]
#5 windSpeed
wind_weather_energy  = boxplot(energy_weather_data_outliers$windSpeed)
min_wind_energy      = min(wind_weather_energy$out)
energy_weather_data_outliers = energy_weather_data_outliers[energy_weather_data_outliers$windSpeed <
                                                      min_wind_energy, ]
#6 energy_mean
mean_weather_energy = boxplot(energy_weather_data_outliers$energy_mean)
min_mean_energy     = min(mean_weather_energy$out)
energy_weather_data_outliers = energy_weather_data_outliers[energy_weather_data_outliers$energy_mean <
                                                              min_mean_energy, ]


### Pressure variable I couldn't remove the outliers 
```


```{r}
#clearing up the memory 

rm(wind_weather_energy)
rm(mean_weather_energy)
rm(min_weather_energy)
rm(std_weather_energy)
rm(max_weather_energy)
rm(median_weather_energy)
rm(num_df)

```


## Data Quality check 

```{r}

# Validating the quality of the data

# validate::validator(,smart_meter_london)




```




### Exploratory Data Analysis


###. Summary statistics ( DOUBLE CHECK )

```{r}
# load the the data set
data(energy_weather_data_outliers)

# inspect the dataset
str(energy_weather_data_outliers)

# summary
summary(energy_weather_data_outliers)

# calculate person correlation coefficient for each pair of variables
cor(energy_weather_data_outliers)
```





## Graphical Analysis 
```{rGraph plotting}
#Graph plotting
hist(energy_weather_data_outliers$energy_mean, 
     main="Mean Energy Consumption", xlab="energy_mean cloumn", border="blue",
     col="green", las=1)
  abline(v = median(energy_weather_data_outliers$energy_mean), lty = 2)
  legend('topright', 'median of mean energy consumption', lty = 2, bty = 'n')

hist(energy_weather_data_outliers$energy_median, 
     main="Median Energy Consumption", xlab="energy_median cloumn", border="blue",
     col="green", las=1)
  abline(v = median(energy_weather_data_outliers$energy_median), lty = 2)
  legend('topright', 'median of median energy consumption', lty = 2, bty = 'n')

hist(energy_weather_data_outliers$energy_std, 
     main="Standard deviation of Energy Consumption", xlab="energy_std cloumn", border="blue",
     col="green", las=1)
  abline(v = median(energy_weather_data_outliers$energy_std), lty = 2)
  legend('topright', 'median of SD energy consumption', lty = 2, bty = 'n')
  
hist(energy_weather_data_outliers$energy_count, 
     main="Count of Energy Consumption", xlab="energy_count cloumn", border="blue",
     col="green", las=1)
  abline(v = median(energy_weather_data_outliers$energy_count), lty = 2)
  legend('topright', 'median of median energy consumption', lty = 2, bty = 'n')
  
hist(energy_weather_data_outliers$energy_max, 
     main="Maximum Energy Consumption", xlab="energy_max cloumn", border="blue",
     col="green", las=1)
  abline(v = median(energy_weather_data_outliers$energy_max), lty = 2)
  legend('topright', 'maximum of median energy consumption', lty = 2, bty = 'n')
  
hist(energy_weather_data_outliers$energy_sum,
     main="Sum of Energy Consumption", xlab="energy_sum cloumn", border="blue",
     col="green", las=1)
  abline(v = median(energy_weather_data_outliers$energy_sum), lty = 2)
  legend('topright', 'Sum of median energy consumption', lty = 2, bty = 'n')
  
hist(energy_weather_data_outliers$energy_sum,
     main="Sum of Energy Consumption", xlab="energy_sum cloumn", border="blue",
     col="green", las=1)
  abline(v = median(energy_weather_data_outliers$energy_sum), lty = 2)
  legend('topright', 'Sum of median energy consumption', lty = 2, bty = 'n')
  
hist(energy_weather_data_outliers$energy_min,
     main="Min of Energy Consumption", xlab="energy_min cloumn", border="blue",
     col="green", las=1)
  abline(v = median(energy_weather_data_outliers$energy_min), lty = 2)
  legend('topright', 'Minimum of median energy consumption', lty = 2, bty = 'n')
  
hist(energy_weather_data_outliers$windSpeed,
     main="Wind Speed during the day", xlab="windSpeed cloumn", border="blue",
     col="green", las=1)
  abline(v = median(energy_weather_data_outliers$windSpeed), lty = 2)
  legend('topright', 'Wind speed median of the the day', lty = 2, bty = 'n')
  
hist(energy_weather_data_outliers$pressure,
     main="Atmospheric pressure during the day", xlab="pressure cloumn", border="blue",
     col="green", las=1)
  abline(v = median(energy_weather_data_outliers$pressure), lty = 2)
  legend('topright', 'Pressure median of the the day', lty = 2, bty = 'n')

hist(energy_weather_data_outliers$humidity,
     main="Humidity during the day", xlab="humidity cloumn", border="blue",
     col="green", las=1)
  abline(v = median(energy_weather_data_outliers$humidity), lty = 2)
  legend('topright', 'Humidity median of the the day', lty = 2, bty = 'n')

hist(energy_weather_data_outliers$temperatureMax,
     main="Maximum temperature during the day", xlab="maximum temperature cloumn", border="blue",
     col="green", las=1)
  abline(v = median(energy_weather_data_outliers$temperatureMax), lty = 2)
  legend('topright', 'Maximum temperature median of the the day', lty = 2, bty = 'n')
  
hist(energy_weather_data_outliers$temperatureLow,
     main="Low temperature during the day", xlab="low temperature cloumn", border="blue",
     col="green", las=1)
  abline(v = median(energy_weather_data_outliers$temperatureLow), lty = 2)
  legend('topright', 'Low temperature median of the the day', lty = 2, bty = 'n')

hist(energy_weather_data_outliers$temperatureMin,
     main="Minimum temperature during the day", xlab="minimum temperature cloumn", border="blue",
     col="green", las=1)
  abline(v = median(energy_weather_data_outliers$temperatureMin), lty = 2)
  legend('topright', 'Maximum temperature median of the the day', lty = 2, bty = 'n')

hist(energy_weather_data_outliers$temperatureHigh,
     main="High temperature during the day", xlab="high temperature cloumn", border="blue",
     col="green", las=1)
  abline(v = median(energy_weather_data_outliers$temperatureHigh), lty = 2)
  legend('topright', 'Maximum temperature median of the the day', lty = 2, bty = 'n')
  
hist(energy_weather_data_outliers$moonPhase,
     main="Moon phase", xlab="high temperature cloumn", border="blue",
     col="green", las=1)
  abline(v = median(energy_weather_data_outliers$temperatureHigh), lty = 2)
  legend('topright', 'Maximum temperature median of the the day', lty = 2, bty = 'n')
```



## Word Cloud part  ( REWORD the #)

```{r}
# setting seed for the reproducibility of the word cloud

set.seed(20)

# Collapsing spaces and storing the values inside a variable

words = paste(energy_weather_data_outliers$summary, collapse = " ")

# Removing superfluous words

words = stringr::str_replace_all(words,"throughout","")
words = stringr::str_replace_all(words,"mostly","")
words = stringr::str_replace_all(words,"partly","")

# Wordcloud code

wordcloud(words = words, min.freq = 1,
          max.words=1000, random.order=FALSE, rot.per=0.35, 
          colors=brewer.pal(8, "Dark2"))

```

# Removing unnecessary components that will not be used in the EDA

energy_weather_data_outliers$summary = NULL
```{r}
# changing the name of dataframe and removing it from the memory
new_energy_weather <- energy_weather_data_outliers
rm(energy_weather_data_outliers)
```


## Correlation Map
*this tells us how many numerical variables are highly correlated (anything above +-0.75 we should double check)*

```{r}
# Create Correlation map of variables

corr.set <- select_if(new_energy_weather,is.numeric)
corr.matrix = round(cor(corr.set),2)
corr.matrix 

# Function

get_upper_tri <- function(corr.matrix){
    corr.matrix[lower.tri(corr.matrix)]= NA
    return(corr.matrix)
}

# using upper triangle function to return the upper half of the correlation map

upper_tri <- get_upper_tri(corr.matrix)
upper_tri

# Upper Triangle( heatmap)

melt_cormat <- melt(upper_tri, na.rm = TRUE)

# Creating the Heatmap

ggheatmap = ggplot(data = melt_cormat, aes(Var2, Var1, fill = value))+
 geom_tile(color = "white")+
 scale_fill_gradient2(low = "blue", high = "red", mid = "white", 
   midpoint = 0, limit = c(-1,1), space = "Lab", 
   name="Pearson\nCorrelation") +
  theme_minimal()+ 
 theme(axis.text.x = element_text(angle = 45, vjust = 1, 
    size = 12, hjust = 1))+
 coord_fixed()

# Adding the coefficients onto the heatmap
corelcoef = ggheatmap + geom_text(aes(Var2, Var1, label = value), color = "black", size = 2) +
theme( 
  axis.title.x = element_blank(),
  axis.title.y = element_blank(),
  panel.grid.major = element_blank(), 
  panel.border = element_blank(),
  panel.background = element_blank(),
  axis.ticks = element_blank(),
  legend.justification = c(1, 0),
  legend.position = c(0.6, 0.7),
  legend.direction = "horizontal")+ guides(fill = guide_colorbar(barwidth = 7, barheight = 1, 
                                                                 title.position = "top", 
                                                                 title.hjust = 0.5))
# Print 
corelcoef
# Removing from memory
rm(numericset)

```

energy mean and sum have a correlation of 1!!

## Further removal of variables

```{r}
# Removing highly correlated variables 
                                                  
energy_weather321 = select(new_energy_weather, -c(moonPhase,
                                                  temperatureMax,
                                                  energy_std,
                                                  dewPoint))
# memory clean up 
rm(corr.set)
rm(ggheatmap)
rm(melt_cormat)
rm(upper_tri)
rm(corr.matrix)
rm(corelcoef)

```
## Preparing the data for PCA 
```{r}
## For PCA all variables need to be numerical
str(energy_weather321)

numeric.df = energy_weather321 %>% 
  select(visibility, windBearing, 
         windSpeed,pressure, 
         humidity, temperatureLow,
         temperatureMin,temperatureHigh,
         energy_median,energy_mean, energy_max,
         energy_count, energy_sum, energy_min)

# Catergorical variables that have been removed prior to running PC
# temperatureLowTime
# temperatureHighTime
# time
# summary
# LCLid

```





( COPIED ALESSANDROS BUT I THINK WE NEED TO CHANGE A FEW THINGS)
## Principal Component Analysis    

```{r} 
#  principle components

# perform PCA on the swiss dataset
#   note: variables are centered and scaled before analysis
pca_energy <- prcomp(numeric.df, center = T, scale. = T)

# inspect the attributes of the PCA object returned by prcomp
attributes(pca_energy)
# see value section of the help for the prcomp for more details
help(prcomp)
```

#### 4. Visual analysis of PCA results{#Visual_analysis_PCA}
```{r}
# calculate the proportion of exaplained variance (PEV) from the std values
pc_energy_var <- pca_energy$sdev^2
pc_energy_var
pc_energy_PEV <- pc_energy_var / sum(pc_energy_var)
pc_energy_PEV

# plot the variance per PC
#   note: this can be done using the plot function on the prcomp object
plot(pca_energy)

# plot the cumulative value of PEV for increasing number of additional PCs
#   note: add an 80% threshold line to inform the feature extraction
#     according to the plot the first 3 PCs should be selected
opar <- par()
plot(
  cumsum(pc_energy_PEV),
  ylim = c(0,1),
  xlab = 'PC',
  ylab = 'cumulative PEV',
  pch = 20,
  col = 'orange' 
)
abline(h = 0.8, col = 'red', lty = 'dashed')
par(opar)

# get and inspect the loadings for each PC
#   note: loadings are reported as a rotation matrix (see lecture)
pc_energy_loadings <- pca_energy$rotation
pc_energy_loadings

# plot the loadings for the first three PCs as a barplot
#   note: two vectors for colours and labels are created for convenience
#     for details on the other parameters see the help for barplot and legend
opar <- par()
colvector = c('red', 'orange', 'yellow', 'green', 'cyan', 'blue')
labvector = c('PC1', 'PC2', 'PC3')
barplot(
  pc_energy_loadings[,c(1:3)],
  beside = T,
  yaxt = 'n',
  names.arg = labvector,
  col = colvector,
  ylim = c(-1,1),
  border = 'white',
  ylab = 'loadings'
)
axis(2, seq(-1,1,0.1))
legend(
  'bottomright',
  bty = 'n',
  col = colvector,
  pch = 15,
  row.names(pc_energy_loadings)
)
par(opar)

# generate a biplot for each pair of important PCs (and show them on the same page)
#   note: the option choices is used to select the PCs - default is 1:2
opar = par()
par(mfrow = c(2,2))
biplot(
  pca_energy,
  scale = 0,
  col = c('grey40','orange')
)
biplot(
  pca_energy,
  choices = c(1,3),
  scale = 0,
  col = c('grey40','orange')
)
biplot(
  pca_energy,
  choices = c(2,3),
  scale = 0,
  col = c('grey40','orange')
)
par(opar)

# the space of the first three PCs is better explored interactively...
#   ...using a function from the pca3d package
# first install pca3d
if(require(pca3d) == FALSE){
    install.packages('pca3d')
}
# then plot and explore the data by rotating/zoom with the mouse
pca3d::pca3d(pca_energy, show.labels = T)

# and save a snapshot of the view in png format
pca3d::snapshotPCA3d('pc_energy_3D.png')
```
















